Model to study co-evolution of trade an culture

### 2015-03-12 article.tex
finish first round of correction on intro

\WSCpagesetup{Montanier, Carrignon, and Zerr}

@@

%XRC trade networks are increasingly being recognised as a key process in ancient Mediterranean
The trade networks are being increasingly recognised as a key process in the structure of ancient Mediterranean societies. Among the recent works studying these networks, brugmans 2010 shows the relation between trade network and the use of table ware in the roman empire. In another example Malkin 2003 show the influence of these trade networks on the emergence of the greek identity between XXX and XXX. 
%An objective of the research undertaken is to compare these networks between each other and understand their specificities. On top of these comparisons we would also like to determine if these trade networks are comparable to a market economy. 
The work presented in this paper takes place within the EPNET project~\cite{remesal_epnet_2014} which aims at studying the properties of the trade network of the roman empire.

%XRC this process played a role in the trajectories of cultural change of the different regions, which were interconnected by trade
The trade possibilities between different regions will play a role in changes in their cultures. The influence can comes from two main sources. First, with each trade connection new good can arrive in a province. The arrival of these new goods will change the rate of adoption of other goods CITE. Moreover the arrival of new products can lead to their recombination with previous products and therefore the creation of an new item. Persons are also travelling on these trade networks, each carrying its own culture. Large movements of populations may lead to changes in the cultures. For example, when Roman troops were travelling to a new region, traditional roman products were delivered to that regions such as oil CITE ROMAN DELIVERING OIL.

%XRC these two elements (cultural transmission and trade networks) are traditionally been studied as isolated systems.
Traditionally cultural transmission and trade are studied distinctively.  On the one hand, the transmission of culture is studied purely as learning and transmitting information between persons or animals. The focus is then for example to understand the different transmission, selection and transformation mechanism. This does not take into account the economical constraints of certain behaviours in human societies. On the other hand, the economical behaviours of persons are studied solely with regards to economical values. The questions traditionally answered are relative to the convergence of prices to an optimal value or the possibility for a person to maximise the economical value of its possessions. The influence of the culture surrounding the person and the bias that may result from it are not taken into account.

%XRC this is wrong because blablabla...
As has been stated before the trade of goods is influence the culture by modifying both the goods present in a region and the people consuming these goods. Moreover we can also expect that the different cultures present in a region will influence the trades this region has with others. \jm{include trade pottery} Focusing on either trade or cultural transmission allows to formalize their inner mechanisms and understand their respective dynamics. However, in order to understand the analyse the trading system of a region one must take into account both trading strategy and culture. \jm{find a problem that we can study thanks to that}
%XRC A potential reviewer would say that you need to specify your research question and needs before justifying the tool
%Jm   -> take care that it is well justified
%XRC we propose a model of coevolution to explore the deep links between trade and cultural transmission in the ancient Mediterranean
%JM we think that agent stuff is a good tool because blah blah
We propose in this article a framework to study the interaction between cultural transmission and trade. This framework should be able first to reproduce well known dynamics of both cultural transmission and trading. Moreover, to be useful, this framework should be able to show the result of the interaction of these dynamics.

%What is agent-based modeling and what can it do
The Agent Based Modeling (ABM) methodology~\cite{} has been applied to both the study of cultural transmission and the study of trading~\cite{}. \jm{achievement on each aspect} ABM is therefore a good candidate to produce a framework able to express theories on trade and cultural transmission. Within these models the social self-organisation of a society results from the interactions of a large number of agents. The behaviour of these agents is encoded so as to reflect the theory under test.
%What we want to bring to ABM with our work
The aim of this paper is to develop an ABM to model the evolution of culture and trade. Importantly we aim to create a model with a low number of parameters in order to ensure that the dynamics observed are simple to analyse. In order to assess the quality of this model we use it to reproduce qualitatively certain dynamics observed in either economy or culture. We will then show the capacities of our model to study the interaction between culture and trade.

>>>>

\WSCpagesetup{Montanier, Carrignon, Rubio and Zerr}

@@

%XRC trade networks are increasingly being recognised as a key process in ancient Mediterranean
Trades are being increasingly recognised as a key process in the structure of ancient Mediterranean societies. For example, the use of table ware in the Roman empire  has been linked to the trade network~\cite{brughmans_connecting_2010}. Interestingly this trade network has been shown to be a market economy~\cite{temin_market_2001}. The work presented in this paper proposes a framework to further study the relation between trade and culture within the roman empire. This work takes place within the EPNET project~\cite{remesal_epnet_2014}.

%XRC this process played a role in the trajectories of cultural change of the different regions, which were interconnected by trade
The trade possibilities between provinces modify the transmission of culture within each province and between them. The influence can comes from two main sources. First, when a new connection is made between two provinces, new products can arrive in their respective markets. The arrival of new goods naturally changes the rate of adoption of every goods of the market CITE. Moreover, the arrival of new products can lead to their recombination with previous products and therefore the creation of an new item. This further modify the structure of the market of the province. Second, persons are also travelling on the roads of the trade network, each carrying its own culture. Large movements of populations may lead to changes in the cultures. For example, when Roman soldiers were stationing in a province, Roman products such as olive oil were delivered to that province CITE ROMAN DELIVERING OIL. This can lead to the adoption of olive oil in regions that were originally not using it.

%XRC these two elements (cultural transmission and trade networks) are traditionally been studied as isolated systems.
Traditionally cultural transmission and trade are studied distinctively.  On the one hand, the cultural transmission mechanisms studied do not take economical consideration into accounts. The works done are interested in the adoption of various cultural trait without taking into consideration their economical value. On the other hand, trading behaviours are studied within their own framework which are designed to measure the amount of goods exchanged and the value at which they are exchanged. Within these frameworks the cultural transmission between persons is not present.

%XRC this is wrong because blablabla
As it has been stated before, trades influence cultural transmission by modifying both the goods present in a region and the people consuming these goods. On the one side, analysing separately trade and cultural transmission allows to formalize their inner mechanisms and understand their respective dynamics. On the other hand, in order to analyse the adoption of goods within a province one must take into account both trading strategy and cultural transmission. For example, the modification of good's consumption after the rupture of a connection in the trade system is better explained by knowing the local adoption of this good (cultural transmission) and the prices of this good (trade).

%XRC we propose a model of coevolution to explore the deep links between trade and cultural transmission in the ancient Mediterranean
We propose in this article a framework to study the interaction between cultural transmission and trade. This framework should be able first to reproduce well known dynamics of both cultural transmission and trading. Second, to be useful, this framework should be able to compare results obtained by taking cultural or trade into consideration. In order to build this framework we view trade events as part of the culture. The price of a good are then viewed as a cultural trait. This traits can be transmitted and modified along different rules. By this mean, traditional works previously expressed in a trading framework, can be expressed in a cultural evolution framework. Therefore, the dynamics observed on various cultural transmission mechanisms can be straightforwardly compared to dynamics obtained by trading mechanisms. In order to demonstrate the feasibility of this approach we propose a minimal implementation able to reproduce both cultural transmission dynamics and trade dynamics. Additionally we will analyse the results obtained with tools from both economy and cultural evolution.
